---
title: "Final results, abstract"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
  word_document: default
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# A Prefix nulling hook.
## see http://stackoverflow.com/questions/22524822/how-can-i-remove-the-prefix-index-indicator-1-in-knitr-output
# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})

knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

```{r, "load packages and data and source files"}
source ("scripts/01_read_data.R")
forest_plot <- readRDS("data/Final forest plot stratified and interaction.RDS")
temp_can_delete <- new.env()
# sa1 <- new.env()
# sa2 <- new.env()
load(file = "data/Data.Rdata", envir = temp_can_delete)
study <- temp_can_delete$study
rm(temp_can_delete)
library(ggplot2)
library(Gmisc)
library(htmlTable)
library(gridExtra)
library(tidyverse)
library(stringr)

```

# Results section
```{r "MACE results"}
load("model_summaries/jags_samples_main.Rdata")
ratio_upper_lower <- format(round(inter_q[3,] / inter_q[2,],2),nsmall = 2)
inter_q <- as.data.frame(t(format(round(inter_q,2), nsmall = 2)))
myrownames <- c("Stratified", "Pooled", "Fixed", "Pooled_tx_strat", "Fixed_tx_strat", "Lau", "Lau plus ours")
row.names(inter_q) <- myrownames
inter_q$interaction <- paste0( inter_q[,1], " (95% CI ", inter_q[,2], " to ", inter_q[,3], ")")
pooled <- inter_q["Pooled" , "interaction"]

men_q <- format(men_q, nsmall = 2)
ratio_upper_lower <- format(round(women_q[3,] / women_q[2,],2),nsmall = 2)
women_q <- as.data.frame(t(format(round(women_q,2), nsmall = 2)))
row.names(women_q) <- c("Three trials", myrownames)
women_q$women <- paste0( women_q[,1], " (95% CI ", women_q[,2], " to ", women_q[,3], ")")
women_q$ratio_upper_lower <- ratio_upper_lower
women_pooled <- women_q[ "Pooled", "women"]

women_q <- format(women_q, nsmall = 2)

rr_prob_r <- readRDS("data/rr_prob_re.RDS")
rr_prob_r <- round(rr_prob_r, 2)
```


## Meta-analysis

### Sex-treatment interaction

```{r, "information criteria"}
load(file = "data/mace_dic.Rdata")
mace_dic <- mace_dic[c("fixed_effects", "random_tx_strat", "random_effects"),]
mace_smry <- lapply(mace_dic, function (x) paste(
  paste(x[-3], collapse = ", "),
  x[3], sep = " and "))
```

On comparing the three models of sex-treatment interactions across all 9 trials of P2Y~12~ inhibitors (Figure 2), the model assuming a shared sex-treatment interaction for the whole drug-class had both a lower deviance and DIC than the other models (the deviance was `r mace_smry$Deviance` and the DIC was `r mace_smry$DIC` for the identical, stratified and shared models respectively), indicating better fit. Consequently this model was preferred.

For this shared-effects model the rate ratio (RR) for the sex-treatment interaction was `r inter_q["Pooled", "50%"]` (95% Credible Interval (CI) `r inter_q["Pooled", "2.5%"]` to `r inter_q["Pooled", "97.5%"]`, Figure 3), indicating approximately 10% lower efficacy in women. Similar estimates for the sex-treatment interactions were found for the identical-effects (RR `r inter_q["Fixed", "50%"]`; 95% CI `r inter_q["Fixed", "2.5%"]` to `r inter_q["Fixed", "97.5%"]`) and independent-effects (RR for prasugrel/ticagrelor versus clopidogrel comparison `r inter_q["Pooled_tx_strat", "50%"]`; 95% CI `r inter_q["Pooled_tx_strat", "2.5%"]` to `r inter_q["Pooled_tx_strat", "97.5%"]`) models.

The treatment effect RR in men for the 3 trials comparing prasugrel or ticagrelor to clopidogrel in ACS was  `r men_q[1]` (95% CI `r men_q[2]` to `r men_q[3]`). Combining this estimate with the sex-treatment interaction estimate from the shared model yielded a RR in women of `r women_q["Pooled", "50%"]` (95% CI `r women_q["Pooled", "2.5%"]` to `r women_q["Pooled", "97.5%"]`, `r women_p["re"]`).


```{r, "bleeding"}
load(file = "model_summaries/bleeding.Rdata")

qfx <- function (x) {
  x <- format( round(x,2), nsmall = 2, trim = TRUE)
  paste0(x[1], " (95% CI ", x[2], " to ", x[3], ")")
}

inter_bleed_q_smr <- qfx(inter_bleed_q)
men_bleed_q_smr <- qfx(men_bleed_q)
women_bleed_q_smr <- qfx(women_bleed_q)

bleed_events_women <- bleed %>% 
  select(starts_with("rw")) %>% 
  as.matrix() %>% 
  sum()

```


For bleeding, the sex-treatment interaction RR was `r inter_bleed_q_smr`. This estimate did not change importantly from the moderately informative prior we used in the modelling for this outcome (prior for sex-treatment interaction on rate ratio scale 1.00; 95% CI 0.80 to 1.25). In men the treatment effect on bleeding was RR `r men_bleed_q_smr`. As such, the modelled treatment effect on bleeding in women was RR `r women_bleed_q_smr`.

### Sex-specific absolute treatment effects in the Scottish Population

```{r, "baseline risk"}
br <- read.csv("data/baseline_risk_actual_data.csv", as.is = TRUE)
br <- br [, names(br) != "surv"]
names(br) <- c("Age", "Sex", "Population", "Bleeding",  "Cardiovascular", "Non-bleeding, non-cardiovascular")
br <- br [ , c( "Sex", "Age", "Population", "Bleeding",  "Cardiovascular", "Non-bleeding, non-cardiovascular")]
br$Age <- gsub("-", " to ", br$Age,fixed = TRUE)
br_gnd <- br %>%
  group_by(Sex) %>%
  summarise (Age = "Total", Population = sum(Population), Bleeding = sum(Bleeding), Cardiovascular = sum(Cardiovascular),
             `Non-bleeding, non-cardiovascular` = sum (`Non-bleeding, non-cardiovascular`))
br_tot <- br %>%
  summarise (Age = "", Sex = "Both", Population = sum(Population), Bleeding = sum(Bleeding), Cardiovascular = sum(Cardiovascular), `Non-bleeding, non-cardiovascular` = sum (`Non-bleeding, non-cardiovascular`))
br <- rbind(br, br_gnd)
br <- br[ order(br$Sex, br$Age),]
br <- rbind(br, br_tot)

br$`All cause` <- br$Bleeding + br$Cardiovascular + br$`Non-bleeding, non-cardiovascular`

br [ , c("Bleeding",  "Cardiovascular", "Non-bleeding, non-cardiovascular", "All cause")] <- 
  lapply (br[, c("Bleeding",  "Cardiovascular", "Non-bleeding, non-cardiovascular", "All cause")], function (x){
    x_p <- round(100* x/ br$Population, 1)
    x_p <- format(x_p, digits = 2, trim = TRUE)
    paste0(x, " (", x_p, "%)")
  })

## Retain labels to allow index referencing for abstract and results
br_reference <- br %>% 
  filter(Age == "Total")

## Remove duplicate labels for tables
br$Sex [duplicated(br$Sex)] <- ""
br$Age[br$Age == "ZZZ"] <- ""

```

17,842 women and 27,818 men were admitted to hospital with myocardial infarction. Mortality was higher for women than men for all-cause mortality (`r br_reference[br_reference$Sex == "Female", "All cause"]` versus `r br_reference[br_reference$Sex == "Male", "All cause"]`) and cardiovascular mortality (`r br_reference[br_reference$Sex == "Female", "Cardiovascular"]` versus `r br_reference[br_reference$Sex == "Male", "Cardiovascular"]`). Mortality related to bleeding was also commoner in women(`r br_reference[br_reference$Sex == "Female", "Bleeding"]` versus `r br_reference[br_reference$Sex == "Male", "Bleeding"]`) deaths. The sex differences in cardiovascular mortality differed by age with younger women having higher mortality than men, and older women having lower mortality than men (Table 2, Figure 4).

```{r, "absolute risk reduction"}
load(file = "data/Results_scotland_effects.Rdata")
arr_prob <- readRDS(file = "data/prob_arr_men_grt_arr_women.Rds")
arr_prob <- round(arr_prob,2)

# re fornmat so include % for main analysis
arr_smry_0avd_non_cllps <- apply(arr_smry_0avd_non_cllps, 1:2, function (x) {
  str_replace_all(x, fixed(" (95% CI"), "% (95% CI")
})
# replace ( with ;) fornmat and include % for sensitivity analysis
arr_smry_0null_mace <- apply(arr_smry_0null_mace, 1:2, function (x) {
  x <- str_replace_all(x, fixed(" (95% CI"), "%; 95% CI")
  str_replace_all(x, fixed(")"), "")
})
# replace - with " to " and 
gender_age_spec_0 [ , c("Bleeding", "Cardiovascular", "Other")] <- 
  lapply(gender_age_spec_0[ , c("Bleeding", "Cardiovascular", "Other")], function (x) {
    x <- str_replace(x, "([0-9])(-)([0-9])", "\\1 to \\3")
    str_replace(x, fixed(" ("), "% (95% CI")
  })
```

Based on our model the absolute risk reduction in all-cause mortality at one-year was similar in both sexes. For the 27,818 men admitted to hospital with
myocardial infarction the modelled absolute risk reduction (ARR) was `r arr_smry_0avd_non_cllps["Men", "Cardiovascular"]` for cardiovascular, `r arr_smry_0avd_non_cllps["Men", "Bleeding"]` for bleeding 
and `r arr_smry_0avd_non_cllps["Men", "Overall"]` for all-cause mortality. For the 17,842 women the ARRs were similar,  `r arr_smry_0avd_non_cllps["Women", "Cardiovascular"]` for cardiovascular, `r arr_smry_0avd_non_cllps["Women", "Bleeding"]` for bleeding and `r arr_smry_0avd_non_cllps["Women", "Overall"]` for all-cause mortality. In contrast with the `r inter_p["re"]` probability that relative treatment efficacy is lower in women than men, the probability that the absolute risk reduction for all-cause mortality was lower in women than men was only `r arr_prob`.

In a subsequent analysis, we modelled the absolute treatment effect
assuming that the relative treatment effect was the same in both sexes. Under this alternative model a larger benefit was seen in women (`r arr_smry_0null_mace["Women", "Overall"]`) than in men (`r arr_smry_0null_mace["Men", "Overall"]`).

### Age and sex-specific absolute treatment effects

The same pattern was evident when the absolute treatment effects were
modelled within each age-group (Figure 4). Assuming no age-treatment
interaction, much larger absolute risk reductions were seen in older
people than in younger people. For example the absolute risk reduction was `r gender_age_spec_0[gender_age_spec_0$Age == "30-39" & gender_age_spec_0$Sex == "Men", "Cardiovascular"]` and `r gender_age_spec_0[gender_age_spec_0$Age == "70-79" & gender_age_spec_0$Sex == "Men", "Cardiovascular"]` in men aged 30-39 and 70-79 respectively.

### Sensitivity analyses

```{r}
pt_est_rng <- readRDS("synth_sa_smry/pt_est_rng.Rds")
```

We modelled the absolute treatment effect for a range of different event
rates for bleeding deaths in women and men. We
allowed for bleeding rates up to 10-fold higher in men than observed,
and up to 2-fold higher in women than in men. In all of these analyses,
the point estimate for the overall absolute treatment effect in women remained
consistently around 2% (range `r pt_est_rng`).



# Abstract

Objective
---------

To estimate the absolute treatment effects of newer P2Y~12~ inhibitors (ticagrelor and prasugrel) compared to clopidogrel in men and women with acute coronary syndrome (ACS).

Methods
------
We searched Ovid MEDLINE, EMBASE and the Cochrane Central Register of Controlled Trials (CENTRAL) for randomised controlled trials of oral P2Y~12~ inhibitors for acute stroke or ACS. Age and sex-specific mortality was obtained for all patients admitted to hospital with myocardial infarction in Scotland from 2006 to 2010 (prior to introduction of prasugrel or ticagrelor).

Results
-------
From 9,277 articles, 9 fulfilled our inclusion criteria. Three trials compared newer P2Y~12~ inhibitors to clopidogrel in ACS, in which the treatment rate ratio (RR) for major adverse cardiovascular events in men was `r men_q[1]` (95% CI `r men_q[2]` to `r men_q[3]`). For the same outcome, across all 9 trials the sex-treatment interaction RR was `r inter_q["Pooled", "interaction"]`. Combining these estimates yielded a treatment RR in women of `r women_q["Pooled", "50%"]` (95% CI `r women_q["Pooled", "2.5%"]` to `r women_q["Pooled", "97.5%"]`).

17,842 women and 27,818 men were admitted to hospital with myocardial infarction. Mortality was higher for women than men for all-cause (`r br_reference[br_reference$Sex == "Female", "All cause"]` versus `r br_reference[br_reference$Sex == "Male", "All cause"]`), cardiovascular (`r br_reference[br_reference$Sex == "Female", "Cardiovascular"]` versus `r br_reference[br_reference$Sex == "Male", "Cardiovascular"]`) and bleeding (`r br_reference[br_reference$Sex == "Female", "Bleeding"]` versus `r br_reference[br_reference$Sex == "Male", "Bleeding"]`) deaths.

On applying the sex-specific RRs to this population, the absolute risk reduction for mortality at 1-year was similar for women and men for all-cause (`r arr_smry_0avd_non_cllps["Women", "Overall"]` versus `r arr_smry_0avd_non_cllps["Men", "Overall"]`), cardiovascular (`r arr_smry_0avd_non_cllps["Women", "Cardiovascular"]`) versus `r arr_smry_0avd_non_cllps["Men", "Cardiovascular"]`) and bleeding (`r arr_smry_0avd_non_cllps["Women", "Bleeding"]` versus `r arr_smry_0avd_non_cllps["Men", "Bleeding"]`) deaths.

Conclusion
----------

Newer P2Y~12~ inhibitors may be slightly less efficacious in women than men (around 1.1-fold lower), but the absolute risk reduction is similar in both sexes.

# Tables and Figures

### Figure 2 and 3

See Script 8b. Had to edit in inkscape to add tables.


## Figure 4 Age and sex specific baseline risks and absolute risk reduction

```{r, fig.height= 4, fig.width=10}
load(file = "data/Bline, ARR plots.Rdata")
tiff(filename = "figures/figure3_bline_arr.tiff", units = "in",
       height = 6, width = 14, compression = "lzw", res = 300)
gridExtra::grid.arrange(bline + guides(colour = FALSE) ,
                        arr_single,
                        layout_matrix = matrix(c(1,2), nrow = 1))
dev.off()
```