---
title: "Final results, abstract"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
  html_document: default
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# A Prefix nulling hook.
## see http://stackoverflow.com/questions/22524822/how-can-i-remove-the-prefix-index-indicator-1-in-knitr-output
# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})

knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

```{r, "load packages and data and source files"}
source ("scripts/01_read_data.R")
forest_plot <- readRDS("data/Final forest plot stratified and interaction.RDS")
temp_can_delete <- new.env()
# sa1 <- new.env()
# sa2 <- new.env()
load(file = "data/Data.Rdata", envir = temp_can_delete)
study <- temp_can_delete$study
rm(temp_can_delete)
library(ggplot2)
library(Gmisc)
library(htmlTable)
library(gridExtra)
library(tidyverse)
library(stringr)

```

# Results section
```{r "MACE results"}
load("model_summaries/jags_samples_main.Rdata")
ratio_upper_lower <- format(round(inter_q[3,] / inter_q[2,],2),nsmall = 2)
inter_q <- as.data.frame(t(format(round(inter_q,2), nsmall = 2)))
myrownames <- c("Stratified", "Pooled", "Fixed", "Pooled_tx_strat", "Fixed_tx_strat", "Lau", "Lau plus ours")
row.names(inter_q) <- myrownames
inter_q$interaction <- paste0( inter_q[,1], " (95% CI ", inter_q[,2], " to ", inter_q[,3], ")")
pooled <- inter_q["Pooled" , "interaction"]

men_q <- format(men_q, nsmall = 2)
ratio_upper_lower <- format(round(women_q[3,] / women_q[2,],2),nsmall = 2)
women_q <- as.data.frame(t(format(round(women_q,2), nsmall = 2)))
row.names(women_q) <- c("Three trials", myrownames)
women_q$women <- paste0( women_q[,1], " (95% CI ", women_q[,2], " to ", women_q[,3], ")")
women_q$ratio_upper_lower <- ratio_upper_lower
women_pooled <- women_q[ "Pooled", "women"]

women_q <- format(women_q, nsmall = 2)

rr_prob_r <- readRDS("data/rr_prob_re.RDS")
rr_prob_r <- round(rr_prob_r, 2)
```


## Meta-analysis

### Sex-treatment interaction

```{r, "information criteria"}
load(file = "data/mace_dic.Rdata")
mace_dic <- mace_dic[c("fixed_effects", "random_tx_strat", "random_effects"),]
mace_smry <- lapply(mace_dic, function (x) paste(
  paste(x[-3], collapse = ", "),
  x[3], sep = " and "))
```

On comparing the three models of sex-treatment interactions across all 8 trials of P2Y~12~ inhibitors, the model assuming a shared sex-treatment interaction for the whole drug-class had both a lower deviance and Deviance Information Criterion (DIC) than the other models (the deviance was `r mace_smry$Deviance` and the DIC was `r mace_smry$DIC` for the identical, stratified and shared models respectively), indicating better fit. Consequently this model was preferred.

For this shared-effects model the rate ratio (RR) for the sex-treatment interaction was `r inter_q["Pooled", "50%"]` (95% Credible Interval (CI) `r inter_q["Pooled", "2.5%"]` to `r inter_q["Pooled", "97.5%"]`, Figure 2), indicating approximately 10% lower efficacy in women. Similar estimates for the sex-treatment interactions were found for the identical-effects (RR `r inter_q["Fixed", "50%"]`; 95% CI `r inter_q["Fixed", "2.5%"]` to `r inter_q["Fixed", "97.5%"]`) and independent-effects (RR for prasugrel/ticagrelor versus clopidogrel comparison `r inter_q["Pooled_tx_strat", "50%"]`; 95% CI `r inter_q["Pooled_tx_strat", "2.5%"]` to `r inter_q["Pooled_tx_strat", "97.5%"]`) models. In the case of the independent model, however, where data was not shared between treatment comparisons and indications, the credible interval was considerably wider.

The treatment effect RR in men for the 3 trials comparing prasugrel or ticagrelor to clopidogrel in acute coronary syndrome was  `r men_q[1]` (95% CI `r men_q[2]` to `r men_q[3]`). Combining this estimate with the sex-treatment interaction estimate from the shared model yielded a RR in women of `r women_q["Pooled", "50%"]` (95% CI `r women_q["Pooled", "2.5%"]` to `r women_q["Pooled", "97.5%"]`, `r women_p["re"]`).


```{r, "bleeding"}
load(file = "model_summaries/bleeding.Rdata")

qfx <- function (x) {
  x <- format( round(x,2), nsmall = 2, trim = TRUE)
  paste0(x[1], " (95% CI ", x[2], " to ", x[3], ")")
}

inter_bleed_q_smr <- qfx(inter_bleed_q)
men_bleed_q_smr <- qfx(men_bleed_q)
women_bleed_q_smr <- qfx(women_bleed_q)

bleed_events_women <- bleed %>% 
  select(starts_with("rw")) %>% 
  as.matrix() %>% 
  sum()

```


For bleeding, the sex-treatment interaction RR was `r inter_bleed_q_smr`. This estimate did not change importantly from the moderately informative prior we used in the modelling for this outcome (prior for sex-treatment interaction on rate ratio scale 1.00; 95% CI 0.80 to 1.25). In men the treatment effect on bleeding was RR `r men_bleed_q_smr`. As such, the modelled treatment effect on bleeding was RR `r women_bleed_q_smr`.

### Sex-specific absolute treatment effects in the Scottish Population

```{r, "baseline risk"}
br <- read.csv("data/baseline_risk_actual_data.csv", as.is = TRUE)
br <- br [, names(br) != "surv"]
names(br) <- c("Age", "Sex", "Population", "Bleeding",  "Cardiovascular", "Non-bleeding, non-cardiovascular")
br <- br [ , c( "Sex", "Age", "Population", "Bleeding",  "Cardiovascular", "Non-bleeding, non-cardiovascular")]
br$Age <- gsub("-", " to ", br$Age,fixed = TRUE)
br_gnd <- br %>%
  group_by(Sex) %>%
  summarise (Age = "Total", Population = sum(Population), Bleeding = sum(Bleeding), Cardiovascular = sum(Cardiovascular),
             `Non-bleeding, non-cardiovascular` = sum (`Non-bleeding, non-cardiovascular`))
br_tot <- br %>%
  summarise (Age = "", Sex = "Both", Population = sum(Population), Bleeding = sum(Bleeding), Cardiovascular = sum(Cardiovascular), `Non-bleeding, non-cardiovascular` = sum (`Non-bleeding, non-cardiovascular`))
br <- rbind(br, br_gnd)
br <- br[ order(br$Sex, br$Age),]
br <- rbind(br, br_tot)
br$Sex [duplicated(br$Sex)] <- ""
br$Age[br$Age == "ZZZ"] <- ""

br$`All cause` <- br$Bleeding + br$Cardiovascular + br$`Non-bleeding, non-cardiovascular`

br [ , c("Bleeding",  "Cardiovascular", "Non-bleeding, non-cardiovascular", "All cause")] <- 
  lapply (br[, c("Bleeding",  "Cardiovascular", "Non-bleeding, non-cardiovascular", "All cause")], function (x){
    x_p <- round(100* x/ br$Population, 1)
    x_p <- format(x_p, digits = 2, trim = TRUE)
    paste0(x, " (", x_p, "%)")
  })
```

There were `r br[br$Sex == "Both", "Population"]` patients admitted to hospital in Scotland with
myocardial infarction over the 5-year period, among whom there were
`r br[br$Sex == "Both", "All cause"]` deaths. The majority of deaths were from cardiovascular
causes `r br[br$Sex == "Both", "Cardiovascular"]`, and bleeding was recorded as the cause
of death in `r br[br$Sex == "Both", "Bleeding"]` cases. Overall cardiovascular deaths
were commoner in women, but bleeding deaths were similar in both sexes
(Table 2). The sex differences in cardiovascular mortality differed by
age with younger women having higher mortality than men, and older women
having lower mortality than men.


```{r, "absolute risk reduction"}
load(file = "data/Results_scotland_effects.Rdata")
arr_prob <- readRDS(file = "data/prob_arr_men_grt_arr_women.Rds")
arr_prob <- round(arr_prob,2)

# re fornmat so include % for main analysis
arr_smry_0avd_non_cllps <- apply(arr_smry_0avd_non_cllps, 1:2, function (x) {
  str_replace_all(x, fixed(" (95% CI"), "% (95% CI")
})
# replace ( with ;) fornmat and include % for sensitivity analysis
arr_smry_0null_mace <- apply(arr_smry_0null_mace, 1:2, function (x) {
  x <- str_replace_all(x, fixed(" (95% CI"), "%; 95% CI")
  str_replace_all(x, fixed(")"), "")
})
# replace - with " to " and 
gender_age_spec_0 [ , c("Bleeding", "Cardiovascular", "Other")] <- 
  lapply(gender_age_spec_0[ , c("Bleeding", "Cardiovascular", "Other")], function (x) {
    x <- str_replace(x, "([0-9])(-)([0-9])", "\\1 to \\3")
    str_replace(x, fixed(" ("), "% (95% CI")
  })
```

Based on our model the absolute risk reduction in all-cause mortality at one-year was similar in both sexes. For the 27,818 men admitted to hospital with
myocardial infarction the modelled absolute risk reduction (ARR) was `r arr_smry_0avd_non_cllps["Men", "Cardiovascular"]`
for cardiovascular, `r arr_smry_0avd_non_cllps["Men", "Bleeding"]` for bleeding 
and `r arr_smry_0avd_non_cllps["Men", "Overall"]` for
all-cause mortality. For the 17,842 women the ARRs were similar,  `r arr_smry_0avd_non_cllps["Women", "Cardiovascular"]` for
cardiovascular, `r arr_smry_0avd_non_cllps["Women", "Bleeding"]` for bleeding and
`r arr_smry_0avd_non_cllps["Women", "Overall"]` for all-cause
mortality. In contrast with the `r inter_p["re"]` probability that relative treatment efficacy is lower in women than men, the probability that the absolute risk reduction for all-cause mortality was lower in women than men was only `r arr_prob`.

In a subsequent analysis, we examined the absolute treatment effect
assuming that the relative treatment effect was the same in both sexes. Under this alternative model a larger benefit was seen in women (`r arr_smry_0null_mace["Women", "Overall"]`) than in men (`r arr_smry_0null_mace["Men", "Overall"]`).

### Age and sex-specific absolute treatment effects

The same pattern was evident when the absolute treatment effects were
modelled within each age-group (Figure 3). Assuming no age-treatment
interaction, much larger absolute risk reductions were seen in older
people than in younger people. For example the absolute risk reduction was `r gender_age_spec_0[gender_age_spec_0$Age == "30-39" & gender_age_spec_0$Sex == "Men", "Cardiovascular"]` and `r gender_age_spec_0[gender_age_spec_0$Age == "70-79" & gender_age_spec_0$Sex == "Men", "Cardiovascular"]` in men aged 30-39 and 70-79 respectively.

### Sensitivity analyses

We modelled the absolute treatment effect for a range of different event
rates for cardiovascular and bleeding deaths in women and men. We
allowed for bleeding rates up to 10-fold higher in men than observed,
and up to 2-fold higher in women than in men. In all of these analyses,
the point estimate for the absolute treatment effect in women remained
consistently above the null (supplementary appendix).



# Abstract

Objective
---------

To estimate the absolute treatment effects of newer P2Y~12~ inhibitors
(ticagrelor and prasugrel) compared to clopidogrel in men and women with
acute coronary syndrome.

Design
------

Systematic review and meta-analysis of randomised clinical trials, and
modelling study with routine healthcare data.

Data Sources
------------

Ovid MEDLINE and EMBASE and the Cochrane Central Register of Controlled
Trials (CENTRAL). National register of all hospitalisations in Scotland
from 2006 to 2010 (prior to introduction of prasugrel or ticagrelor).

Eligibility criteria
--------------------

Randomised controlled trials of oral P2Y~12~ inhibitors for stroke or
acute coronary syndrome where sex-specific estimates were provided.
Trials of intravenous P2Y~12~ inhibitors or that that randomised patients to different durations
of P2Y~12~ therapy were excluded.

Modelled outcomes
-----------------

Cardiovascular, bleeding and all-cause mortality.

Results
-------

From 7,148 articles, 36 were reviewed in detail with 8 fulfilling our
inclusion criteria. Trials compared clopdigorel to placebo in acute coronary syndrome (3 trials) and stroke (2 trials) or newer P2Y~12~ inhibitors to clopidogrel in acute coronary syndrome (3 trials). This provided a total of 109,570 randomised participants. For all 8 trials, for major adverse cardiovascular events the sex-treatment interaction rate ratio (RR) was `r inter_q["Pooled", "interaction"]`, such that the probability that relative treatment efficacy for P2Y~12~ inhibitors is lower in women than in men is `r rr_prob_r`.

For the 3 trials of prasugrel or ticagrelor compared to clopidogrel for acute coronary syndrome, the treatment RR  in men was `r men_q[1]` (95% CI `r men_q[2]` to `r men_q[3]`). Combining this estimate with the sex-treatment interaction for the drug-class yielded a treatment RR in women of `r women_q["Pooled", "50%"]` (95% CI `r women_q["Pooled", "2.5%"]` to `r women_q["Pooled", "97.5%"]`).

On applying these (and major bleeding) treatment effect estimates to mortality data for Scotland; For the 27,818 men admitted to hospital with myocardial infarction the modelled absolute risk reduction (ARR) at one year was `r arr_smry_0avd_non_cllps["Men", "Cardiovascular"]` for cardiovascular, `r arr_smry_0avd_non_cllps["Men", "Bleeding"]` for bleeding and `r arr_smry_0avd_non_cllps["Men", "Overall"]` for all-cause mortality. For the 17,842 women the ARRs were similar, `r arr_smry_0avd_non_cllps["Women", "Cardiovascular"]` for cardiovascular, `r arr_smry_0avd_non_cllps["Women", "Bleeding"]` for bleeding and `r arr_smry_0avd_non_cllps["Women", "Overall"]` for all-cause mortality, such that the probability that the absolute risk reduction for all-cause mortality was lower in women than men was only `r arr_prob`.

Conclusion
----------

In a meta-analysis and modelling study using data from 100,000 participants randomised to P2Y~12~ inhibitors and over 45,000 patients from the Scottish population hospitalized with myocardial infarction we have identified that newer P2Y~12~ inhibitors may be slightly less efficacious in women than men (around 1.1-fold lower), but that the absolute risk reduction is similar in both sexes

# Tables and Figures

### Figure 2 Events and rates for MACE with figure
```{r, fig.height=7, fig.width=14}
forest_plot <- readRDS(file = "data/Final forest plot stratified and interaction.RDS")
forest_plot
```


## Figure 3 Age and sex specific baseline risks and absolute risk reduction

```{r, fig.height= 4, fig.width=10}
load(file = "data/Bline, ARR plots.Rdata")
tiff(filename = "figures/figure3_bline_arr.tiff", units = "in",
       height = 6, width = 14, compression = "lzw", res = 300)
gridExtra::grid.arrange(bline + guides(colour = FALSE) ,
                        arr_single +
                          theme(axis.text.y = element_blank(), axis.title.y = element_blank()),
                        layout_matrix = matrix(c(1,2), nrow = 1))
dev.off()
```